---
title: "Leaflet legend cutting of"
author: "Chris Madsen"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = F)

library(tidyverse)
library(ggspatial)
library(bcdata)
library(sf)
library(leaflet)
library(rasterbc)
```

```{r options, include = F}
dist_from_wb_to_search = 10 # This number is in kilometers

buffer_size_in_m = 50

elevation_threshold = 10 # meters.

slope_threshold = 5 # this number is in degrees.

invasive_species = 'smallmouth_bass'

invasive_species_label = stringr::str_to_title(stringr::str_replace(invasive_species,'_',' '))

focus_wb = 'Echo Lake'

focus_wb_code_name = stringr::str_to_lower(stringr::str_replace(focus_wb, ' ','_'))

focus_graph_name = paste0(stringr::str_extract(focus_wb_code_name,'^[a-z]+'),'_graph')

out_folder_name = paste0(invasive_species,'_',dist_from_wb_to_search,'km_radius_',buffer_size_in_m,'_stream_buffer_',elevation_threshold,'_elev_',slope_threshold,'_slope')

if(!dir.exists(paste0('output/',out_folder_name))){
  dir.create(paste0('output/',out_folder_name))
}

file.remove(list.files(path = paste0('output/',out_folder_name), full.names = T))
```

# `r focus_wb` - Waterbody Connectivity

## Parameters of Assessment

- This run is for `r invasive_species_label`, with `r focus_wb` as the source waterbody.
- The search radius around `r focus_wb` is `r dist_from_wb_to_search` km.
- Dispersal of the species is deemed possible if the change in elevation AND/OR slope between two waterbodies is less than or equal to `r elevation_threshold` / `r slope_threshold` respectively.
- Waterbodies are buffered by `r buffer_size_in_m` to facilitate identifying overlaps that otherwise might not be identified due to slight errors in stream/river/lake polygon extents.

```{r which_wb_is_ours}
# ws = read_sf("W:/CMadsen/shared_data_sets/WatershedGroups_lowres.shp") |> select(WATERSHE_1)
# 
# echo_lakes = bcdc_query_geodata('freshwater-atlas-lakes') |>
#   filter(GNIS_NAME_1 == 'Echo Lake') |>
#   collect()

# # Join to watersheds to see which is on vancouver island.
# 
# echo_lakes |>
#   st_join(ws, st_intersects) |> View()
```

```{r get_wb_polygon}
focus_wb_poly = bcdc_query_geodata('freshwater-atlas-lakes') |> 
  filter(GNIS_NAME_1 == focus_wb,
         WATERSHED_GROUP_ID == 20) |> 
  collect()
```

## `r focus_wb`

### First Glance {.tabset}

#### Leaflet map

```{r}
centroid_points = as.data.frame(st_centroid(focus_wb_poly) |> st_transform(crs = 4326) |> st_coordinates())

leaflet() |> 
  addTiles() |> 
  addPolygons(
    fillColor = 'darkblue',
    label = ~GNIS_NAME_1,
    data = focus_wb_poly |> sf::st_transform(4326)
  ) |> 
  addScaleBar('bottomright') |> 
  setView(lng = centroid_points$X, lat = centroid_points$Y, zoom = 14)
```

#### Nearby lakes, rivers and streams

```{r get_watershed_atlas}

# Get all lake/stream/river polygons within X kilometers, depending on the
# options we wrote above.
nlakes = bcdc_query_geodata('freshwater-atlas-lakes') |> 
  filter(DWITHIN(focus_wb_poly, dist_from_wb_to_search, 'kilometers')) |> collect()

nstreams = bcdc_query_geodata('freshwater-atlas-stream-network') |>
  filter(DWITHIN(focus_wb_poly, dist_from_wb_to_search, 'kilometers')) |> 
  # filter(FEATURE_SOURCE != 'lake-def skelet') |> 
  collect() 

# # Remove stream lines that are actually lake defining skeletons 
# # or have 'River' in the name.
# nstreams = nstreams |>
#   filter(FEATURE_SOURCE != 'lake-def skelet') |> 
#   filter(!stringr::str_detect(GNIS_NAME,'River$'))

nrivers = bcdc_query_geodata('freshwater-atlas-rivers') |>
  filter(DWITHIN(focus_wb_poly, dist_from_wb_to_search, 'kilometers')) |> collect()
```

```{r simplify columns and join together into graphs.}
nlakes = nlakes |> 
  dplyr::mutate(polytype = 'lake') |> 
  dplyr::select(name = GNIS_NAME_1, polytype)

nrivers = nrivers |> 
  dplyr::mutate(polytype = 'river') |> 
  dplyr::select(name = GNIS_NAME_1, polytype)

nstreams = nstreams |> 
  dplyr::mutate(polytype = 'stream') |> 
  dplyr::select(name = GNIS_NAME, polytype)

# Buffer everything by 10 meters so that slight overlaps are almost guaranteed - we need these to carry out a spatial union.
nlakes = st_buffer(nlakes, buffer_size_in_m)
nrivers = st_buffer(nrivers, buffer_size_in_m)
nstreams = st_buffer(nstreams, buffer_size_in_m)

wbs = nlakes |> 
  bind_rows(nrivers, nstreams)

wbs = wbs |>
  dplyr::select(name,
                polytype)

### Leaflet map of all waterbodies within `r dist_from_echo_lake_to_search` km of `r focus_wb`
```

```{r leaflet_of_all_wbs}

ids_for_leaf = paste0(wbs$name, " (",wbs$polytype,')')
ids_for_leaf = str_replace_all(ids_for_leaf, "^NA", "nameless")

leaflet() |> 
  addTiles() |> 
  addPolygons(
    color = ~case_when(name == focus_wb ~ 'red',
                       is.na(name) ~ 'blue',
                       T ~ 'blue'),
    fillColor = ~case_when(name == focus_wb ~ 'red',
                           is.na(name) ~ 'blue',
                           T ~ 'blue'),
    fillOpacity = ~case_when(name == focus_wb ~ 1,
                             is.na(name) ~ 0.5,
                             T ~ 0.5),
    label = ids_for_leaf,
    data = wbs |> st_transform(crs = 4326)
  ) |> 
  addScaleBar('bottomright')
```

### Waterbody 'Graph' Maps {.tabset}

#### Without Elevation/Slope

```{r leaflet_of_graphs}
parts = st_cast(st_union(wbs),"POLYGON")

clust = unlist(st_intersects(wbs, parts))

wbs$graph_id = clust

wbs = st_transform(wbs, 4326)

# wbs |> 
#   st_drop_geometry() |> 
#   count(graph_id, sort = T)
# Mostly in graph 1.
# Let's put all graphs past graph ID 6 as "Other"

wbs_f = wbs |> 
  mutate(graph_id = as.factor(graph_id)) |> 
  mutate(graph_id = forcats::fct_lump(graph_id, n = 6))

wbs_s = wbs_f |> 
  group_by(graph_id) |> 
  summarise()

my_pal = colorFactor(palette = 'Spectral',
                     domain = unique(wbs_f$graph_id))

leaflet() |>
  addProviderTiles(providers$CartoDB) |>
  addPolygons(
    fillOpacity = 0,
    color = 'yellow',
    opacity = 1,
    weight = 5,
    label = ~name,
    data = nngeo::st_remove_holes(wbs |> filter(name == focus_wb))
  ) |> 
  addPolygons(
    color = 'black',
    weight = 1.5,
    fillColor = ~my_pal(graph_id),
    fillOpacity = 0.8,
    label = ~paste0("Graph: ",graph_id),
    data = wbs_s
  ) |>
  addScaleBar('bottomright') |> 
  addLegend(pal = my_pal,
            title = 'Graph ID',
            values = wbs_s$graph_id)
```

```{r adding_elevation_to_leaflet, include=F,message=F,results='hide'}
# Get elevation for this area.
my_bbox = st_as_sfc(st_bbox(wbs_s))

# rasterbc::listdata_bc()
# datadir_bc('C:/Users/CMADSEN/Downloads/LocalR/echo_lake_bass/data/', quiet=TRUE)

# getdata_bc(geo=my_bbox, collection='dem', varname='slope')

# terra::rast(tif.path)

my_elevation <- elevatr::get_elev_raster(locations = my_bbox, z = 9, clip = "locations")

my_slope = terra::terrain(my_elevation, opt = 'slope', units = 'degrees')

# Not using flow yet. But could be useful!
my_flow_direction = terra::terrain(my_elevation, 'flowdir')

# elev_pal = colorNumeric(c("#331b0e", "#359c3d", "#93ccc5"), terra::values(my_elevation),
#   na.color = "transparent")


```

#### With Elevation/Slope

```{r leaflet_of_slope_and_elevation}
elev_pal = colorNumeric(c("darkgreen", "#359c3d", "yellow","red","white"), terra::values(my_elevation),
  na.color = "transparent")

slope_pal <- colorNumeric(c("#000340", "#bdc0ff"), terra::values(my_slope),
  na.color = "transparent")

leaflet() |> 
  addProviderTiles(providers$CartoDB, group = 'CartoDB') |> 
  addPolygons(
    color = 'black',
    weight = 2,
    fillColor = ~my_pal(graph_id),
    fillOpacity = 0.6,
    label = ~paste0("Graph: ",graph_id),
    data = wbs_s
  ) |> 
    addRasterImage(group = 'elevation', my_elevation, colors = elev_pal, opacity = 0.4) |>
    addRasterImage(group = 'slope', my_slope, colors = slope_pal, opacity = 0.8) |> 
  addLayersControl(baseGroups = 'CartoDB',
                   overlayGroups = c("elevation","slope"),
                  options = layersControlOptions(collapsed = F)) |> 
  addPolygons(
    fillColor = 'green',
    fillOpacity = 0.8,
    color = 'green',
    label = ~name,
    data = wbs |> filter(name == focus_wb)
  )
```
